---
title: "PH241 Final Project - Model Fits"
output: html_notebook
---

***
### Setup
```{r, warning=FALSE, message=FALSE}
# Importing Libraries
library(dplyr)
library(data.table)

# Setting File Paths
prefix = "~/PH241/Final Project/Data/Cleaned/"
setwd(prefix)
filepaths = c("washb-bangladesh-merged-public.csv")

mergedData = fread(filepaths[1], data.table=FALSE) %>% select(-V1)
```

Let's begin with an overview of the data we'll be looking at. Our merged dataset includes one binary outcome variable (diar7d) and a variety of explanatory variables. Each row represents a measurement of a child's diarrheal status over the last 7 days, where 0 indicates no diarrhea and 1 indicates some report of diarrhea, where diarrhea was defined as >= 3 loose or watery stools in 24 hours or >= 1 stool with blood over the 7 days prior to measurement.

```{r}
mergedData %>% head
```

Now, let's fit some models.

### Model 1 - Control Arm
In examining the control arm, we want to know whether "better" sanitary and hygienic conditions among the nonrandomized controls contributes to decreased diarrheal disease. If there is a difference, is that difference due to confounding? 

```{r}
controlData = mergedData %>% filter(tr=="Control")
controlData %>% nrow
controlData %>% head
```


